Eye in a Disk: eyeIntegration human pan-eye and body transcriptome database version 1.0

### 2019-06-08 manuscript.Rmd
fixed vinny comments

To ensure there are no substantial differences in quantification of the gene TPM, we calculated the $R^2$ between the GTEX-calculated TPM values and our length scaled and transformed TPM values we calculate for our GTEx samples (see methods). We got an $R^2$ of 0.89. Zhang et al. report that RNA-seq quantifications done between alignment-free methods (used in EiaD) and alignment-based methods (used by GTEx) get a pearson correlation ranging from 0.89 to 0.93 [@zhangEvaluationComparisonComputational2017]. As Zhang et al. compared quantification methods with identical gene references (we use Gencode GRCh38 gene models and GTEx uses hg19) and did not scale TPM score differently, our result falls in line with expectations.

@@

```{r appScreenshots, echo = F, fig.cap = appScreenshots_cap, fig.width=12}
appScreenshots_cap <- fig_cap(name='appScreenshots', caption = 'Screenshots from eyeIntegration web app. A. Pan-tissue gene expression box plots with accompanying data tables. The data tables display the rank (lower is more highly expressed) of each gene in each sub tissue, the decile of the rank (10 is the highest decile of expression), and the mean gene\'s log2(TPM + 1) score for each sub tissue. B. Heatmap visualization.')

@@

Salmon quantification, while highly performant and accurate, has a higher variance when with lower read depth and with shorter transcripts [@zhangEvaluationComparisonComputational2017]. Extra care should be taken with comparisons with lower counts of samples (cornea, RGC) as smaller sample numbers decrease the confidence in differential expression. We recommend that if you wish to directly compare your RNA-seq sample quantifications with EiaD, you should not directly compare our TPM values with yours as there are many important variables that will differ. Instead run our Snakemake pipeline (https://www.github.com/davemcg/EiaD_build) and add your samples. Finally, we would like to remind any users that RNA-seq methods measure mRNA levels, but the functional unit is the protein; westerns are still the gold standard with which to evaluate expression and localization.

## Data accessibility 
Individual data files for gene expression and sample metadata can be downloaded from the ‘Data’ tab on the web app. All data and code used to generate the web app can be installed from the R command line by running devtools::install_github(‘davidmcg/eyeIntegration_app’). The code for the EiaD data processing pipeline can be found at https://github.com/davidmcg/Eyeintegration_autobuild.

>>>>

To ensure there are no substantial differences in quantification of gene TPM values, we determined the $R^2$ between GTEX and EiaD generated TPM values for our shared GTEx samples (see methods); we computed an $R^2$ of 0.89. “ Zhang et al. report that RNA-seq quantifications done between alignment-free methods (used in EiaD) and alignment-based methods (used by GTEx) get a $R^2$ ranging from 0.89 to 0.93 [@zhangEvaluationComparisonComputational2017]. As Zhang et al. compared quantification methods with identical gene references (we use Gencode GRCh38 gene models and GTEx uses hg19) and did not scale TPM score differently, our result falls in line with expectations.

@@

```{r appScreenshots, echo = F, fig.cap = appScreenshots_cap, fig.width=12}
appScreenshots_cap <- fig_cap(name='appScreenshots', caption = 'Screenshots from eyeIntegration web app. A. Pan-tissue gene expression box plots with accompanying data tables. The data tables display the rank (lower is more highly expressed) of each gene in each sub tissue, the decile of the rank (10 is the highest decile of expression), and the gene\'s mean log2(TPM + 1) score for each sub tissue. B. Heatmap visualization.')

@@

Salmon quantification, while highly performant and accurate, has a higher variance for lower read depth samples and shorter transcripts [@zhangEvaluationComparisonComputational2017]. Extra care should be taken with comparisons with lower counts of samples (cornea, RGC) as smaller sample numbers decrease the confidence in differential expression. We do not recommend you directly compare our TPM values with your counts data as there are many important variables that will differ. Instead run our Snakemake pipeline (https://www.github.com/davemcg/EiaD_build), adding your samples. Finally, we would like to remind any users that RNA-seq methods measure mRNA levels, but the functional unit is the protein; westerns are still the gold standard with which to evaluate expression and localization.

## Data accessibility 
Individual data files for gene expression and sample metadata can be downloaded from the ‘Data’ tab on the web app. All data and code used to generate the web app can be installed from the R command line by running devtools::install_github(‘davidmcg/eyeIntegration_app’). The code for the EiaD data processing pipeline can be found at https://github.com/davemcg/EiaD_build.
